---
title: "Untitled"
output: pdf_document
date: "2026-03-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(car)
library(alr4)
library(leaps)
library(countrycode)


transit_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-05/transit_cost.csv')

```


```{r data cleaning}

clean_transit <- transit_cost %>%
  select(c(e, country, city, start_year, end_year, rr, length, tunnel_per, tunnel, stations, year, ppp_rate, real_cost, cost_km_millions)) %>%
  filter(!is.na(e) & real_cost != 0)

clean_transit <- clean_transit %>%
  mutate(real_cost = as.numeric(real_cost)) %>%
  mutate(rr = as.factor(rr)) %>%
  filter(end_year <= 2026) %>%
  mutate(duration = as.numeric(end_year) - as.numeric(start_year))

```



```{r region mapping}

clean_transit %>%
  group_by(country) %>%
  count(sort = T)

gb <- filter(clean_transit, clean_transit$country == "UK") %>%
  mutate(country = "GB")

clean_transit<- subset(clean_transit, !e %in% c(7226, 7336, 7417))
clean_transit <- rbind(clean_transit, gb)
  

clean_transit$region <- countrycode(sourcevar = clean_transit$country,
                            origin = "iso2c",
                            destination = "region23")

df <- clean_transit %>%
  select(c(e, rr, length, stations, ppp_rate, real_cost, cost_km_millions, duration, region, year))

clean_transit %>%
  group_by(region) %>%
  summarize(mean_cost = mean(real_cost), median_cost = median(real_cost), mean_duration = mean(duration, na.rm=T), mean_length = mean(length), mean_cost_by_km = mean(cost_km_millions)) 

```


```{r region_merge}

clean_transit <- clean_transit %>%
  mutate(group = case_when(
    region %in% c("Northern America", "Central America", "South America") ~ "Americas",
    region %in% c("Western Europe", "Northern Europe", "Southern Europe", "Eastern Europe", "Northern Africa") ~ "Europe & North Africa",
    region %in% c("Eastern Asia", "Southern Asia", "Central Asia", "Western Asia") ~ "Asia",
    region %in% c("South-Eastern Asia", "Australia and New Zealand") ~ "SE Asia & Oceania"
  ))

df <- clean_transit
unique(clean_transit["group"])
unique(df$group)
table(df$group)
```


## Data Exploration 


### Fit full model
```{r}
full <- lm(real_cost ~ group + year+ length + stations+ ppp_rate+duration, data = df)
summary(full)
vif(full)
plot(full) 
plot(full_log) 
hist(full$residuals)


```

It turns out that we need do some transformation because the variance is non-constant and the distribution of standardized residuals is not normal. 



### Transformation- Use BoxCox

```{r}
library(MASS)
df_pos <- df %>% filter(real_cost > 0)
full_pos <- lm(real_cost ~ region + year + length + stations + ppp_rate + duration, data = df_pos)
MASS::boxcox(full_pos, plotit = TRUE) ### The result turns out that we should raise y to power of 0.25
```
```{r}
full_trans <- lm(real_cost^0.25 ~ group + year+ length + stations+ ppp_rate+duration, data = df)
summary(full_trans)
vif(full_trans)
hist(full_trans$residuals)
plot(full_trans) 
```





### Check for outliers

```{r}
outlierTest(full_log)
```

### Model selection 
#### Apply forward, backward, and bidirection
```{r}
reg <- regsubsets(
  real_cost ~ group + year+ length + stations+ ppp_rate+duration,
  data = df,
  nbest = 1,
  nvmax = 13
)
reg_summary <- summary(reg)
bic_min_index <- which.min(reg_summary$bic)
bic_min_model <- reg_summary$which[bic_min_index, ]
cp_min_index <- which.min(reg_summary$cp)
cp_min_model <- reg_summary$which[cp_min_index, ]
bic_min_model
cp_min_model
```
BIC: Exclude duration
Cp: Exclude duration + ppp_rate



```{r}
step(full_trans,
     direction = 'backward')
```

Choose not to remove any predictors



