```{r setup, include = FALSE}
library(tidyverse)
library(sf)
library(spdep)
library(car)
library(alr4)
library(leaps)
library(gridExtra)
library(countrycode)
library(dplyr)
library(broom)
transit_cost <- read.csv("transitcost.csv")
transit_cost$tunnel_per = as.numeric(gsub("[\\%,]", "", transit_cost$tunnel_per))
culture <- read.csv("hofstede_iso.csv")
colnames(culture)[[1]] <- "ISO"
```

```{r data cleaning, include = FALSE}

clean_transit <- transit_cost %>%
  select(c(e, country, city, start_year, end_year, rr, length, 
  tunnel_per, tunnel, stations, year, ppp_rate, real_cost,
   cost_km_millions)) %>%
  filter(!is.na(e) & real_cost != 0)

clean_transit <- clean_transit %>%
  mutate(real_cost = as.numeric(real_cost)) %>%
  mutate(rr = as.factor(rr)) %>%
  filter(end_year <= 2026) %>%
  mutate(duration = as.numeric(end_year) - as.numeric(start_year))

```

```{r region mapping}

gb <- filter(clean_transit, clean_transit$country == "UK") %>%
  mutate(country = "GB")

clean_transit<- subset(clean_transit, !e %in% c(7226, 7336, 7417))
clean_transit <- rbind(clean_transit, gb)
  

clean_transit$region <- countrycode(sourcevar = clean_transit$country,
                            origin = "iso2c",
                            destination = "region23")

```

```{r region_merge}

clean_transit <- clean_transit %>%
  mutate(group = case_when(
    region %in% c("Northern America", "Central America", "South America") ~ "Americas",
    region %in% c("Southern Europe", "Northern Europe", "Western Europe") ~ "Europe",
    region %in% c("Central Asia", "Eastern Asia") ~ "Asia",
    region %in% c("South-Eastern Asia", "Australia and New Zealand") ~ "SE Asia & Oceania",
    region %in% c("Northern Africa", "Western Asia") ~ "North Africa & West Asia",
    .default = region
  ))



#table(clean_transit$group)
df <- clean_transit
df <- df[!is.na(df$tunnel_per), ]
```


# Transit Costs across the World

## Data

Our transit data comes from the Transit Costs Project collaboration at the NYU Marron Institute of Urban Management. The express purpose of this dataset is to document the costs of transit projects across the world for international comparison. The transit projects under consideration all involve a railway, including light rail, trams, rapid transit/metro systems, and heavy rail (the meaning of this terminology changes based on local vernacular). This means busses, ferries, rideshare, bikes or other forms of transportation are not included in the dataset. This data was compiled by academic researchers and documents transit projects from . Each observation is a unique transit project undertaken by that country 

To account for spatial factors, country shapefiles from a public ArcGIS dataset were merged into the . The map projection used in the dataset is the Mercator projection, which emphasizes accuracy of shape over size. The final dataset that was merged with the transit dataset was a cross-cultural psychology dataset from social psychologist Geert Hofstede. This dataset includes Hofstede's rankings of countries along six cultural dimensions: individualism-collectivism, uncertainty avoidance, strength of social hierarchy (called 'power-distance'), motivation toward achievement and success (formerly masculine v. feminine), and long-term vs. short term orientation. The observations in the Hofstede dataset were individual countries (or groupings of countries such as "Arab states").

## Data Cleaning

We remove a number of variables such as `source` and `currency` that serve only to tell us the source of the data or the local currency, so we can work with variables relevant to our questions such as `real_cost` and `length`. We then set the `real_cost` variable to be a number, and filter out observations with missing variables or that have yet to be completed (as we do not care for projects yet to begin or yet to finish). 

We create three new variables:
`duration` (dbl): how long the project took, computed as the difference between the end and start years
`region` (character): which geographic region the transit project's country is located in, as per the country's ISO code
`group` (character): a further grouping of projects by geographic regions, to ensure that there are sufficient observations to represent different geographic areas

## Introduction

### Project Goal
To analyze and compare transit projects around the world to understand the factors that affect the cost and speeds of projects around the world, including the additional impact of geographic or cultural factors.

### Research Questions
#### 1. Project Speed
What variables are significant when predicting the speed at which transit projects are completed? 

#### 2. Project Cost
Which variables are significant when predicting the cost of a transit project? 

#### 3. Spatial and Cultural factors
Are spatial and cultural factors significant for models that predict the cost and speed of a project? 

## Methodology 

Three models were fit to the combined data.

### Model 1 - Project Speed

$$
Y_i = \beta_0 + \beta_1 \cdot X_{1i} + \beta_2 \cdot X_{2i} + \beta_3 \cdot X_{3i} + \sum_{j = 4}^{9} \beta_j \cdot X_{ji}
$$

Where:  
$X_{1i}$ = the natural log of the real cost of the project in millions of USD
$X_{2i}$ = length of the project in km
$X_{3i}$ = indicator of whether the project is based on railroad or not (0 if not, 1 if yes)
And the sum notes the effect of the regional group the project it belongs to. There are 7 regional groups, with the base case being the Americas

### Model 2 - Cost Prediction
$$
Y_i = \beta_0 + \beta_1X_{1i} + \beta_2X_{2i} + \beta_3X_{3i} + \beta_4X_{4i} + 
$$

### Model 3 -

### Model 4 - 

## Results


### Project Speed (Sangey Rinchen)

With this model, we are interested in answering the question of whether or not a model can predict how quickly a transit project can be completed, and which predictors are required to answer that question. There isn't a specific variable for that in the dataset, so we begin by creating the variables we need and begin creating our model.  

#### Data cleaning and the full model

We need a variable for the project speed, which we calculate with the simple formula:  

$$\frac{\text{length}}{\text{duration}}$$:  

```{r speedData}

df_speed <- df %>%
  filter(!is.na(duration)) %>%
  mutate(project_speed = length / duration)

rownames(df_speed) <- NULL #resetting row names

```

We begin with our initial model, factoring in possible explanatory variables (a few such as year and duration are excluded, and attempt to find the best subset using BIC, adjusted R squared, and Mallow's $C_p$. Note that we use the natural log of the real cost to avoid any possible issues that may arise due to how much bigger the numbers in this variable are compared to the rest of the dataset.    

```{r fullModel}

subsets <- summary(regsubsets(project_speed ~ log(real_cost) + rr + length + 
                      stations + cost_km_millions + group, nbest = 1, nvmax = 8, data = df_speed))
subsets$adjr2
subsets$cp
subsets$bic
subsets$outmat

```

Though there isn't one lone model all metrics agree on, subset 6 has satisfactory scores and will be used (though with all groups in the `group` variable present).  

#### Building the model

```{r speedModel}

speedMdl <- lm(project_speed ~ log(real_cost) + rr + length + stations + group, data = df_speed)
summary(speedMdl)
plot(speedMdl)

#test for multicollinearity
vif(speedMdl)

```

Apart from some potential outliers and a bit of skewing on the Q-Q plots, there isn't anything too concerning. However, we observe some collinearity between `length` and `stations`, understandable given that the two can be used to understand the size of the project in their own ways. Thus, we will remove the `stations` variable and diagnose our model.  

```{r speedMdlUpdates}

speedMdl2 <- lm(project_speed ~ log(real_cost) + rr + length + group, data = df_speed)
summary(speedMdl2)
plot(speedMdl2, which = c(1, 5))
vif(speedMdl2)

```

#### Outliers

Our updated model doesn't appear to have issues with multicollinearity. The residuals plot looks fine but we see some points with high leverage and high residuals. Let's take a closer look at these points to determine whether they should be removed as outliers.  

```{r outlierCheck}

outlierTest(speedMdl2)
df_speed[c(231, 377, 383, 384),]

```

These observations have some of the fastest `project_speed`, longest `length`, and largest `real_cost` of any of the transit projects in the dataset. We will note these outliers, and remove them from our model for now.  

```{r outlierUpdate}

speedMdl3 <- lm(project_speed ~ log(real_cost) + rr + length + group, data = df_speed[-c(231, 377, 383,384),])
plot(speedMdl3, which = 5)
outlierTest(speedMdl3)

#We still see outliers, so we clear them out one more time

df_speed2 <- df_speed[-c(231, 377, 383,384),]

speedMdl4 <- lm(project_speed ~ log(real_cost) + rr + length + group, data = df_speed2[-c(98, 102, 302, 379, 343),])
plot(speedMdl4, which = c(2,5))

#Nothing glaring in the residuals vs leverage plot, so we're done with outliers. However, the Q-Q plot is skewed. Let's fix that:  

boxcox(speedMdl4)

```

#### Final model
```{r finalspeedModel}

df_speed3 <- df_speed2 %>%
  mutate(speed = log(project_speed))
speedMdl_fin <- lm(speed ~ log(real_cost) + rr + length + group, data = df_speed3[-c(98, 102, 302, 379, 343),])

#Last checks:

plot(speedMdl_fin)
vif(speedMdl_fin)
outlierTest(speedMdl_fin)

```

And so, our final model's coefficients are as follows:  
```{r coefficients}

speedMdl_fin$coefficients

```

On observation, it makes sense that the project would take longer if the lines involved setting up railroads (hence `rr1` has a negative coefficient). We observe a typically faster project completion speed in projects in Asia , and a slower completion speed in Europe.  

### Cost Prediction ()

#### Fit the baseline linear model
We model **real construction cost** as a function of (i) geographic regions and
(ii) key characteristics that are plausibly associated with cost: project year, length of line,
number of stations, purchasing power parity (PPP), and construction duration.

$$
real\_cost \sim region + year + length + stations + ppp\_rate + duration
$$

The initial model statistics is

```{r}
full <- lm(real_cost ~ group + year+ length + stations+ ppp_rate+duration, data = df)

broom::glance(full) %>%
  dplyr::select(r.squared, sigma, AIC, BIC, nobs) %>%
  knitr::kable(digits = 3, caption = "Model fit statistics")

```
Linear regression relies on zero means, constant error variance, and approximately normal residuals. We evaluate these using standard residual diagnostics. In the baseline model, residual plots indicate heteroskedasticity and departures from normality.



#### Diagnostic assessment of modeling assumptions


```{r baseline_diagnostics, fig.height=3, fig.width=5}
par(mfrow = c(2,2), mar=c(3,3,2,1))
plot(full, which = 1)
plot(full, which = 2)
plot(full, which = 4)
hist(resid(full), main = "Residuals", xlab = "")
par(mfrow = c(1,1))
```





#### Transformation via BoxCox
To address non-constant variance and improve residual normality, we apply a Box–Cox transformation
to the strictly positive response.


```{r boxcox, fig.width=4, fig.height=3}
library(MASS)
df_pos <- df %>% filter(real_cost > 0)
full_pos <- lm(real_cost ~ group + year + length + stations + ppp_rate + duration, data = df_pos)
par(mar = c(3,3,2,1))
MASS::boxcox(full_pos, plotit = TRUE) ### The result turns out that we should raise y to power of 0.25
```

The Box–Cox log-likelihood plot suggests a transformation near $\lambda \approx 0.25$, so we proceed with $Y^{0.25}$

#### Refit our model and check conditions again. 


```{r residual plot,fig.width=5, fig.height=3}
full_trans <- lm(real_cost^0.25 ~ group + year+ length + stations+ ppp_rate+duration, data = df)
#summary(full_trans)

par(mfrow = c(2,2), mar=c(3,3,2,1))
plot(full_trans, which = 1)
plot(full_trans, which = 2)
plot(full_trans, which = 4)
hist(resid(full_trans), main = "Residuals", xlab = "")
par(mfrow = c(1,1))

```


After transformation, the residual diagnostics improve noticeably. The TA plot suggests more constant variance, and the Q–Q plot shows residuals closer to normality.

We also examine influential observations. Although some points have larger standardized residuals, influence depends on both residual size and leverage, summarized by Cook’s distance. All Cook’s distances are well below common thresholds, 0.5, so no observations are removed.


#### Multicollinearity assessment
```{r VIF, fig.width=3, fig.height=2}
v <- vif(full_trans)

vif_df <- data.frame(
  variable = rownames(v),
  vif = v[,3]
)


ggplot(vif_df, aes(x = vif, y = reorder(variable, vif)))+
  geom_point(size = 3, color = 'darkblue')+
  geom_vline(xintercept = 5, linetype = 'dashed', color = 'red', linewidth=1 )+
  labs(title = 'VIF of variables', y ='variable')+
  theme_classic()
```

We assess multicollinearity using variance inflation factors (VIF). Values exceeding 5 are often used as thresholds to indicate high collinearity. The VIF results here do not indicate problematic collinearity.



#### Model selection 
#### Backward model selection 

To assess whether a simpler model provides similar explanatory power, we apply backward model selection using AIC starting from the transformed-response model. Because AIC balances model fit and complexity, smaller values indicate a better model.

```{r stepwise regression}
step_fit <- step(full_trans,
     direction = 'backward', trace  = 0)

data.frame(
  Initial_AIC = round(AIC(full_trans),2),
  End_AIC = round(AIC(step_fit),2),
  Equation = paste(deparse(formula(step_fit)), collapse = ''))%>%
knitr::kable(caption = "Backward AIC selection summary")

```



**Selection result.** The backward AIC procedure does not remove any predictors, indicating that
each variable contributes meaningfully to model fit under the AIC criterion. It indicates that removing any single predictor increases AIC relative to the full transformed model.



### Final Model

We proceed with the transformed-response model including all pre-specified predictors:

This model achieves improved adherence to the regression assumptions after the transformation, including approximately constant variance and residuals that are closer to normally distributed. Retaining all predictors also preserves interpretability with respect to the study objective, allowing us to evaluate how project characteristics and geographic region jointly influence transit construction costs. Below is the full summary of the model.

```{r}
# coefficient table
broom::tidy(full_trans) %>%
  dplyr::select(term, estimate, std.error, p.value) %>% 
  knitr::kable(digits = 3,caption = "Summary of the transformed regression model.")

# model fit
broom::glance(full_trans) %>%
  dplyr::select(r.squared, adj.r.squared, sigma) %>% 
  knitr::kable(digit = 3)

```

### Spatial and Cultural Predictors (Kevin Ji)

A particular concern for our prior models is spatial autocorrelation. There is strong reason to believe that the presence of transit in neighboring countries could disincentivise or incentivise countries to build it themselves. If a country's passenger rail buildout is successful, it encourages neighbors to invest in rail projects themselves. Certain countries have legal frameworks or cultural biases that discourage infrastructure investments like transit, and these features are often shared among neighboring countries. Spatial autocorrelation may also capture the aspects of geography that make it more costly or inexpensive to build in certain regions, such as soil quality or climate. 

```{r fullmerge, echo = FALSE}

colnames(transit_cost) <- c("e", "ISO", "city", "line",
"start_year", "end_year", "rr", "length", 'tunnel_per',
'tunnel', 'stations', 'source1', 'cost', 'currency',
'year', 'ppp_rate', 'real_cost', 'cost_km_millions',
'source2', 'reference')

```

To model spatial autocorrelation, we import shapefiles from the 

```{r spatial autocorrelation,  out.height="35%", message = FALSE,  warning = FALSE, include = FALSE}

medcost <- transit_cost %>% group_by(ISO) %>% summarize(median.cost = log(median(cost_km_millions, na.rm=T)))

colnames(medcost) <- c("ISO", "Median_Cost")

shp <- st_read("World_Countries_Generalized.shp")

medshp <- merge(shp, medcost)

centroid <- st_centroid(medshp)
dist <- knearneigh(centroid, k = 7)
dist_nb <- knn2nb(dist)
medshp_wts <- nb2listw(dist_nb, style = "W")

```

```{r neighbor plot}
plot(st_geometry(medshp))
plot(dist_nb, coords = st_coordinates(centroid), add = TRUE, col = "red")
```

```{r MC Test, include = FALSE}
moran.test(medshp$Median_Cost, medshp_wts)    

mc <- moran.mc(medshp$Median_Cost, medshp_wts, nsim = 999)

```

```{r Moran Plot, echo = FALSE, message= F}

# Moran Scatterplot

moran.plot(as.numeric(scale(medshp$Median_Cost)), listw=medshp_wts, 
           xlab="Standardized Transit Project Cost", 
           ylab="Standardized Lagged Project Cost",
           main=c("Moran Scatterplot for Transit Costs") )


```

```{r spatiosocial, echo = FALSE}
colnames(clean_transit)[[1]] <- "e"
colnames(clean_transit)[[2]] <- "ISO"
transit.culture <- merge(clean_transit, culture)

transit.culture <- transit.culture %>% 
    mutate(real_cost = as.numeric(real_cost)) %>% 
    mutate(rr = as.factor(rr)) %>% 
    filter(end_year <= 2026) %>%
    mutate(duration = as.numeric(end_year) - as.numeric(start_year)) %>%
    mutate(longtermview = as.numeric(ltowvs),
    uncertainty = as.numeric(uai),
    competitive = as.numeric(mas),
    power_distance = as.numeric(pdi),
    individual = as.numeric(ivr))
    
socio <- transit.culture %>% 
    group_by(ISO) %>%
    summarize(ltv = (abs(median(longtermview))),
    uai = (abs(median(uncertainty) - 100)),
    idv = (median(individual)), 
    pdi =  (median(power_distance)),
    comp = (median(competitive))) %>% 
    ungroup()

socioshp <- merge(shp, socio)

med.p <- ggplot(data = medshp) + 
    geom_sf(aes(fill = Median_Cost)) + theme_void() +
     scale_color_viridis_b(aes = c("fill")) + 
    labs(title = "Median Transit Cost by Country")

p <- ggplot(socioshp) + scale_color_viridis_b(aes = c("fill")) + 
    theme_void()
ltv.p <- p +
    geom_sf(aes(fill = ltv)) +
    labs(title = "Long Term View by Country")

uai.p <- p + 
    geom_sf(aes(fill = uai)) +
    labs(title = "Uncertainty Aversion by Country")

comp.p <- p +
    geom_sf(aes(fill = comp)) +
    labs(title = "Competitiveness by Country")

```
```{r cultural plots, echo = FALSE}
grid.arrange(med.p, ltv.p, uai.p, comp.p, ncol = 2)
```


Some econometric procedures recommending accounting for spatial autocorrelation by inflating standard errors.

```{r Effective N, echo = FALSE}
# effective N for spatial autocorrelation
I <- mc$statistic
eff_n <- nrow(transit_cost)*((1-I)/(1+I))
```

```{r culture model, echo = FALSE, warning = FALSE}


multicheck <- lm(cost_km_millions ~ longtermview + 
uncertainty + competitive + power_distance + 
individual, data = transit.culture)



long_term <- lm(log(cost_km_millions) ~  uncertainty + power_distance + competitive + year + length + ppp_rate + rr + tunnel_per + length + duration + group, data = transit.culture)

vif(long_term)

ncvTest(long_term)
wts <- 1/(abs(rstandard(long_term)))

weighted_socio <- merge(transit.culture, as.data.frame(wts), by = "row.names")

final <- lm(log(cost_km_millions) ~  uncertainty + year + competitive + rr + tunnel_per + length + ppp_rate + duration + group, data = weighted_socio, weights = wts)


```

## Conclusion



