```{r setup, include = FALSE}
library(tidyverse)
library(sf)
library(spdep)
library(car)
library(alr4)
library(leaps)
library(gridExtra)
library(countrycode)
library(dplyr)
library(broom)
transit_cost <- read.csv("transitcost.csv")
transit_cost$tunnel_per = as.numeric(gsub("[\\%,]", "", transit_cost$tunnel_per))
culture <- read.csv("hofstede_iso.csv")
colnames(culture)[[1]] <- "ISO"
```

```{r data cleaning, include = FALSE}

clean_transit <- transit_cost %>%
  select(c(e, country, city, start_year, end_year, rr, length, 
  tunnel_per, tunnel, stations, year, ppp_rate, real_cost,
   cost_km_millions)) %>%
  filter(!is.na(e) & real_cost != 0)

clean_transit <- clean_transit %>%
  mutate(real_cost = as.numeric(real_cost)) %>%
  mutate(rr = as.factor(rr)) %>%
  filter(end_year <= 2026) %>%
  mutate(duration = as.numeric(end_year) - as.numeric(start_year))

```

```{r region mapping}

gb <- filter(clean_transit, clean_transit$country == "UK") %>%
  mutate(country = "GB")

clean_transit<- subset(clean_transit, !e %in% c(7226, 7336, 7417))
clean_transit <- rbind(clean_transit, gb)
  

clean_transit$region <- countrycode(sourcevar = clean_transit$country,
                            origin = "iso2c",
                            destination = "region23")

```

```{r region_merge}

clean_transit <- clean_transit %>%
  mutate(group = case_when(
    region %in% c("Northern America", "Central America", "South America") ~ "Americas",
    region %in% c("Southern Europe", "Northern Europe", "Western Europe") ~ "Europe",
    region %in% c("Central Asia", "Eastern Asia") ~ "Asia",
    region %in% c("South-Eastern Asia", "Australia and New Zealand") ~ "SE Asia & Oceania",
    region %in% c("Northern Africa", "Western Asia") ~ "North Africa & West Asia",
    .default = region
  ))



#table(clean_transit$group)
df <- clean_transit
df <- df[!is.na(df$tunnel_per), ]
```


# Transit Costs across the World

## Data

Our transit data comes from the Transit Costs Project collaboration at the NYU Marron Institute of Urban Management. The express purpose of this dataset is to document the costs of transit projects across the world for international comparison. The transit projects under consideration all involve a railway, including light rail, trams, rapid transit/metro systems, and heavy rail (the meaning of this terminology changes based on local vernacular). This means busses, ferries, rideshare, bikes or other forms of transportation are not included in the dataset. This data was compiled by academic researchers and documents transit projects from . Each observation is a unique transit project undertaken by that country 

To account for spatial factors, country shapefiles from a public ArcGIS dataset were merged into the . The map projection used in the dataset is the Mercator projection, which emphasizes accuracy of shape over size. The final dataset that was merged with the transit dataset was a cross-cultural psychology dataset from social psychologist Geert Hofstede. This dataset includes Hofstede's rankings of countries along six cultural dimensions: individualism-collectivism, uncertainty avoidance, strength of social hierarchy (called 'power-distance'), motivation toward achievement and success (formerly masculine v. feminine), and long-term vs. short term orientation. The observations in the Hofstede dataset were individual countries (or groupings of countries such as "Arab states").



## Methodology 

Three models were fit to the combined data.

### 

### Model 2 - Cost Prediction
$$
Y_i = \beta_0 + \beta_1X_{1i} + \beta_2X_{2i} + \beta_3X_{3i} + \beta_4X_{4i} + 
$$



## Results

### Transit Project Duration

### Cost Prediction ()

#### Fit the baseline linear model
We model **real construction cost** as a function of (i) geographic regions and
(ii) key characteristics that are plausibly associated with cost: project year, length of line,
number of stations, purchasing power parity (PPP), and construction duration.

$$
real\_cost \sim region + year + length + stations + ppp\_rate + duration
$$

The initial model statistics is

```{r}
full <- lm(real_cost ~ group + year+ length + stations+ ppp_rate+duration, data = df)

broom::glance(full) %>%
  dplyr::select(r.squared, sigma, AIC, BIC, nobs) %>%
  knitr::kable(digits = 3, caption = "Model fit statistics")

```
Linear regression relies on zero means, constant error variance, and approximately normal residuals. We evaluate these using standard residual diagnostics. In the baseline model, residual plots indicate heteroskedasticity and departures from normality.



#### Diagnostic assessment of modeling assumptions


```{r baseline_diagnostics, fig.height=3, fig.width=5}
par(mfrow = c(2,2), mar=c(3,3,2,1))
plot(full, which = 1)
plot(full, which = 2)
plot(full, which = 4)
hist(resid(full), main = "Residuals", xlab = "")
par(mfrow = c(1,1))
```





#### Transformation via BoxCox
To address non-constant variance and improve residual normality, we apply a Box–Cox transformation
to the strictly positive response.


```{r boxcox, fig.width=4, fig.height=3}
library(MASS)
df_pos <- df %>% filter(real_cost > 0)
full_pos <- lm(real_cost ~ group + year + length + stations + ppp_rate + duration, data = df_pos)
par(mar = c(3,3,2,1))
MASS::boxcox(full_pos, plotit = TRUE) ### The result turns out that we should raise y to power of 0.25
```

The Box–Cox log-likelihood plot suggests a transformation near $\lambda \approx 0.25$, so we proceed with $Y^{0.25}$

#### Refit our model and check conditions again. 


```{r residual plot,fig.width=5, fig.height=3}
full_trans <- lm(real_cost^0.25 ~ group + year+ length + stations+ ppp_rate+duration, data = df)
#summary(full_trans)

par(mfrow = c(2,2), mar=c(3,3,2,1))
plot(full_trans, which = 1)
plot(full_trans, which = 2)
plot(full_trans, which = 4)
hist(resid(full_trans), main = "Residuals", xlab = "")
par(mfrow = c(1,1))

```


After transformation, the residual diagnostics improve noticeably. The TA plot suggests more constant variance, and the Q–Q plot shows residuals closer to normality.

We also examine influential observations. Although some points have larger standardized residuals, influence depends on both residual size and leverage, summarized by Cook’s distance. All Cook’s distances are well below common thresholds, 0.5, so no observations are removed.


#### Multicollinearity assessment
```{r VIF, fig.width=3, fig.height=2}
v <- vif(full_trans)

vif_df <- data.frame(
  variable = rownames(v),
  vif = v[,3]
)


ggplot(vif_df, aes(x = vif, y = reorder(variable, vif)))+
  geom_point(size = 3, color = 'darkblue')+
  geom_vline(xintercept = 5, linetype = 'dashed', color = 'red', linewidth=1 )+
  labs(title = 'VIF of variables', y ='variable')+
  theme_classic()
```

We assess multicollinearity using variance inflation factors (VIF). Values exceeding 5 are often used as thresholds to indicate high collinearity. The VIF results here do not indicate problematic collinearity.



#### Model selection 
#### Backward model selection 

To assess whether a simpler model provides similar explanatory power, we apply backward model selection using AIC starting from the transformed-response model. Because AIC balances model fit and complexity, smaller values indicate a better model.

```{r stepwise regression}
step_fit <- step(full_trans,
     direction = 'backward', trace  = 0)

data.frame(
  Initial_AIC = round(AIC(full_trans),2),
  End_AIC = round(AIC(step_fit),2),
  Equation = paste(deparse(formula(step_fit)), collapse = ''))%>%
knitr::kable(caption = "Backward AIC selection summary")

```



**Selection result.** The backward AIC procedure does not remove any predictors, indicating that
each variable contributes meaningfully to model fit under the AIC criterion. It indicates that removing any single predictor increases AIC relative to the full transformed model.



### Final Model

We proceed with the transformed-response model including all pre-specified predictors:

This model achieves improved adherence to the regression assumptions after the transformation, including approximately constant variance and residuals that are closer to normally distributed. Retaining all predictors also preserves interpretability with respect to the study objective, allowing us to evaluate how project characteristics and geographic region jointly influence transit construction costs. Below is the full summary of the model.

```{r}
# coefficient table
broom::tidy(full_trans) %>%
  dplyr::select(term, estimate, std.error, p.value) %>% 
  knitr::kable(digits = 3,caption = "Summary of the transformed regression model.")

# model fit
broom::glance(full_trans) %>%
  dplyr::select(r.squared, adj.r.squared, sigma) %>% 
  knitr::kable(digit = 3)

```

### Spatial and Cultural Predictors (Kevin Ji)

A particular concern for our prior models is spatial autocorrelation. There is strong reason to believe that the presence of transit in neighboring countries could disincentivise or incentivise countries to build it themselves. If a country's passenger rail buildout is successful, it encourages neighbors to invest in rail projects themselves. Certain countries have legal frameworks or cultural biases that discourage infrastructure investments like transit, and these features are often shared among neighboring countries. Spatial autocorrelation may also capture the aspects of geography that make it more costly or inexpensive to build in certain regions, such as soil quality or climate. 

```{r fullmerge, echo = FALSE}

colnames(transit_cost) <- c("e", "ISO", "city", "line",
"start_year", "end_year", "rr", "length", 'tunnel_per',
'tunnel', 'stations', 'source1', 'cost', 'currency',
'year', 'ppp_rate', 'real_cost', 'cost_km_millions',
'source2', 'reference')

```

To model spatial autocorrelation, we import shapefiles from the 

```{r spatial autocorrelation,  out.height="35%", message = FALSE,  warning = FALSE, include = FALSE}

medcost <- transit_cost %>% group_by(ISO) %>% summarize(median.cost = log(median(cost_km_millions, na.rm=T)))

colnames(medcost) <- c("ISO", "Median_Cost")

shp <- st_read("World_Countries_Generalized.shp")

medshp <- merge(shp, medcost)

centroid <- st_centroid(medshp)
dist <- knearneigh(centroid, k = 7)
dist_nb <- knn2nb(dist)
medshp_wts <- nb2listw(dist_nb, style = "W")

```

```{r neighbor plot}
plot(st_geometry(medshp))
plot(dist_nb, coords = st_coordinates(centroid), add = TRUE, col = "red")
```

```{r MC Test, include = FALSE}
moran.test(medshp$Median_Cost, medshp_wts)    

mc <- moran.mc(medshp$Median_Cost, medshp_wts, nsim = 999)

```

```{r Moran Plot, echo = FALSE, message= F}

# Moran Scatterplot

moran.plot(as.numeric(scale(medshp$Median_Cost)), listw=medshp_wts, 
           xlab="Standardized Transit Project Cost", 
           ylab="Standardized Lagged Project Cost",
           main=c("Moran Scatterplot for Transit Costs") )


```

```{r spatiosocial, echo = FALSE}
colnames(clean_transit)[[1]] <- "e"
colnames(clean_transit)[[2]] <- "ISO"
transit.culture <- merge(clean_transit, culture)

transit.culture <- transit.culture %>% 
    mutate(real_cost = as.numeric(real_cost)) %>% 
    mutate(rr = as.factor(rr)) %>% 
    filter(end_year <= 2026) %>%
    mutate(duration = as.numeric(end_year) - as.numeric(start_year)) %>%
    mutate(longtermview = as.numeric(ltowvs),
    uncertainty = as.numeric(uai),
    competitive = as.numeric(mas),
    power_distance = as.numeric(pdi),
    individual = as.numeric(ivr))
    
socio <- transit.culture %>% 
    group_by(ISO) %>%
    summarize(ltv = (abs(median(longtermview))),
    uai = (abs(median(uncertainty) - 100)),
    idv = (median(individual)), 
    pdi =  (median(power_distance)),
    comp = (median(competitive))) %>% 
    ungroup()

socioshp <- merge(shp, socio)

med.p <- ggplot(data = medshp) + 
    geom_sf(aes(fill = Median_Cost)) + theme_void() +
     scale_color_viridis_b(aes = c("fill")) + 
    labs(title = "Median Transit Cost by Country")

p <- ggplot(socioshp) + scale_color_viridis_b(aes = c("fill")) + 
    theme_void()
ltv.p <- p +
    geom_sf(aes(fill = ltv)) +
    labs(title = "Long Term View by Country")

uai.p <- p + 
    geom_sf(aes(fill = uai)) +
    labs(title = "Uncertainty Aversion by Country")

comp.p <- p +
    geom_sf(aes(fill = comp)) +
    labs(title = "Competitiveness by Country")

```
```{r cultural plots, echo = FALSE}
grid.arrange(med.p, ltv.p, uai.p, comp.p, ncol = 2)
```


Some econometric procedures recommending accounting for spatial autocorrelation by inflating standard errors.

```{r Effective N, echo = FALSE}
# effective N for spatial autocorrelation
I <- mc$statistic
eff_n <- nrow(transit_cost)*((1-I)/(1+I))
```

```{r culture model, echo = FALSE, warning = FALSE}


multicheck <- lm(cost_km_millions ~ longtermview + 
uncertainty + competitive + power_distance + 
individual, data = transit.culture)



long_term <- lm(log(cost_km_millions) ~  uncertainty + power_distance + competitive + year + length + ppp_rate + rr + tunnel_per + length + duration + group, data = transit.culture)

vif(long_term)

ncvTest(long_term)
wts <- 1/(abs(rstandard(long_term)))

weighted_socio <- merge(transit.culture, as.data.frame(wts), by = "row.names")

final <- lm(log(cost_km_millions) ~  uncertainty + year + competitive + rr + tunnel_per + length + ppp_rate + duration + group, data = weighted_socio, weights = wts)


```

## Conclusion



